\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2023-07-03 cubble.Rmd
section 3.1

A common task in spatio-temporal data analysis is to identify related temporal series from geographically proximate locations. For example, we may want to associate weather stations with nearby river sensors to find the water level series that reflect changes in precipitation. This problem can be generalised by first matching locations across two datasets and then calculating the temporal features of the series. In cubble, locations from two datasets can be matched using the function `match_spatial()`. The function calculates the distance matrix of the two datasets and returns `spatial_n_group` groups with the smallest distances. For each group, analysts can include more location with the argument `spatial_n_each` (default to 1 for one-on-one matching).

@@

The similarity score between two time series is calculated using a matching function, which can be customised by the analysts based on the time series feature relevant to match. The matching function takes two time series as a list and returns a single numerical score. By default, cubble uses a simple peak matching algorithm (`match_peak`) to count the number of peaks in two time series that fall within a specified temporal window.

>>>>

A useful task in spatio-temporal data analysis is to combine related temporal series within a close geographic neighborhood. For example, we may want to examine data from weather stations with water flow records in nearby river sensors to understand how precipitation relates to water supply. This might be useful for better predicting potential for drought and flood.

This problem can be tackled by first matching spatial locations across two datasets and then calculating the temporal features of the multiple series. In cubble, locations from two datasets can be matched using the function `match_spatial()`. The function calculates the distance matrix of the two datasets and returns `spatial_n_group` groups with the smallest distances. For each group, analysts can include more locations with the argument `spatial_n_each` (default to 1 for one-on-one matching).

@@

The similarity score is computed by a matching function which can be customized to any desired time series feature. The matching function takes two time series as a list and returns a single numerical value. By default, cubble uses a simple peak matching algorithm (`match_peak`) to count the number of peaks in two time series that fall within a specified time window.

### 2023-07-03 cubble.tex
section 3.1

Finding nearby locations across two spatio-temporal data sets measuring different processes is a common task in spatio-temporal data analysis. For example, we might want to related measurements obtained from weather stations and river gauge levels recorded at different locations and the goal is to find nearby river sensors that reflect the change in precipitation recorded by the weather station.

This type of problem can be generalize to matched pairs of locations accross different data sets or more generality groups of locations

or groups with a number of locations selected depending on the problem at hand. For example, we might want to find for a given weather station the closest river gauges to compare levels

For example, we might want to find a group of close stations and then use score functions to describe time series features between the series. In cubble, matching locations based on distance can be performed using the function \texttt{match\_spatial()} which finds matched groups in two cubble objects by first fiding the closests pairs and.

Two arguments are available to control the matching: the argument \texttt{spatial\_n\_group} specifies the number of paired groups to return and the argument \texttt{spatial\_n\_each} specifies the number of matches for each item in the first cubble object (default to 1 for one-to-one matching).

and matching time series features for selected locations\texttt{match\_temporal()}. The \texttt{match\_spatial()} function :

\begin{verbatim}
match_spatial(<cubble_obj1>, <cubble_obj2>, ...)
\end{verbatim}

The function \texttt{match\_temporal()} takes the outputs from spatial matching and calculates a similarity score of the time series between spatially matched pairs. The temporal matching requires two identifiers: one for separating each spatially matched group: \texttt{match\_id} and one for separating the two data sources: \texttt{data\_id}. Matching between different variables can be specified using the \texttt{temporal\_by} argument, similar to the \texttt{by} syntax from dplyr's \texttt{*\_join}.

@@

The similarity score between two time series is calculated using a matching function, which can be customised by the analysts based on the time series feature relevant to match. The matching function takes two time series as a list and returns a single numerical score. By default, cubble uses a simple peak matching algorithm (\texttt{match\_peak}) to count the number of peaks in two time series that fall within a specified temporal window.

>>>>

A useful task in spatio-temporal data analysis is to combine related temporal series within a close geographic neighborhood. For example, we may want to examine data from weather stations with water flow records in nearby river sensors to understand how precipitation relates to water supply. This might be useful for better predicting potential for drought and flood.

This problem can be tackled by first matching spatial locations across two datasets and then calculating the temporal features of the multiple series. In cubble, locations from two datasets can be matched using the function \texttt{match\_spatial()}. The function calculates the distance matrix of the two datasets and returns \texttt{spatial\_n\_group} groups with the smallest distances. For each group, analysts can include more locations with the argument \texttt{spatial\_n\_each} (default to 1 for one-on-one matching).

The temporal matching then calculates a similarity score between the time series of spatially matched pairs using the function \texttt{match\_temporal()}. The temporal matching requires two identifiers: one for separating each spatially matched group: \texttt{match\_id} and one for separating the two data sources: \texttt{data\_id}. Matching between different variables can be specified using the \texttt{temporal\_by} argument, similar to the \texttt{by} syntax from dplyr's \texttt{*\_join}.

@@

The similarity score is computed by a matching function which can be customized to any desired time series feature. The matching function takes two time series as a list and returns a single numerical value. By default, cubble uses a simple peak matching algorithm (\texttt{match\_peak}) to count the number of peaks in two time series that fall within a specified time window.
